---
title: "Project Code"
author: "Hugh Chen"
date: "3/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data intro
```{r}
library(hexbin)
mdata = read.csv("movie_rt_transform.csv")
head(mdata)
```

## Exploratory analysis
```{r}
hist(mdata$gross, main="Histogram of Gross", xlab="Gross")
hist(mdata$budget, main="Histogram of Budget", xlab="Budget")
hist(mdata$imdb_score, main="Histogram of IMDB Scores", xlab="IMDB Score")
hist(mdata$rt_score, main="Histogram of Rotten Tomato Scores", xlab="Rotten Tomato Score")

t_rt_score = qnorm((mdata$rt_score+0.5)/101,0)
hist(t_rt_score)

plot(hexbin(mdata$gross, mdata$budget), main="Gross vs. Budget" , legend=0, xlab="Gross (USD)", ylab="Budget (USD)")
plot(hexbin(mdata$imdb_score, mdata$rt_score), main="IMDB Score vs. RT Score" , legend=0, xlab="IMDB Score", ylab="RT Score")
plot(hexbin(log(mdata$gross), log(mdata$budget)), main="Gross vs. Budget" , legend=0, xlab="log gross", ylab="log budget")
```

## Univariate summaries
```{r}
mdata3 = mdata
mdata3$norm_gross = mdata$gross/max(mdata$gross)
mdata3$norm_budget = mdata$budget/max(mdata$budget)
mdata3$norm_imdb_score = mdata$imdb_score/max(mdata$imdb_score)
mdata3$norm_rt_score = mdata$rt_score/max(mdata$rt_score)
boxplot(mdata3[,c(35, 36, 37, 38)], main="Univariate Summary")
```

## Pairwise scatter plots
First the untransformed scatter plot.  
```{r}
mdata2 = mdata
hexplom(~mdata2[,c(2,6,8,30)])
```

Then transform gross and profit by log transform.
```{r}
mdata2$log_gross = log(mdata2$gross)
mdata2$log_budget = log(mdata2$budget)
hexplom(~mdata2[,c(36,35,8,30)])
```

## Residuals for simple models
```{r}
model = lm(mdata$gross~mdata$budget+mdata$imdb_score+mdata$rt_score)
summary(model)
plot(model$fitted, model$resid, main="residual plot", xlab="Fitted", ylab="Residuals")

model = lm(log(mdata$gross)~log(mdata$budget)+mdata$imdb_score+mdata$rt_score)
summary(model)
plot(model$fitted, model$resid, main="residual plot", xlab="Fitted", ylab="Residuals")
```


## Power Transform
Try exact transformation.
```{r}
library(car)
ans <- powerTransform(cbind(mdata$gross, mdata$budget, 
                            mdata$imdb_score, mdata$rt_score+.01) ~ 1)
powerTransform(cbind(mdata$gross, mdata$budget, 
                     mdata$imdb_score, mdata$rt_score+.01) ~ 1)
summary(ans)

gross_trans = (mdata$gross^0.295 - 1)/0.295
budget_trans = (mdata$budget^0.255 - 1)/0.255
imdb_trans = (mdata$imdb_score^2.4 - 1)/2.4
rt_trans = (mdata$rt_score^.79 - 1)/.79
model = lm(gross_trans~budget_trans+imdb_trans+rt_trans)
hexplom(~cbind(gross_trans,budget_trans,imdb_trans,rt_trans))
summary(model)
plot(model$fitted, model$resid, main="residual plot", xlab="Fitted", ylab="Residuals")
```

Try approximate.
```{r}
t_gross = (mdata$gross^0.25 - 1)/0.25
t_budget = (mdata$budget^0.25 - 1)/0.25
t_imdb_1 = mdata$imdb_score
t_imdb_2 = (mdata$imdb_score^2 - 1)/2
t_rt = mdata$rt_score
model1 = lm(t_gross~t_budget+t_imdb_1+t_imdb_2+t_rt)
hexplom(~cbind(t_gross,t_budget,t_imdb_2,t_rt))
summary(model1)
plot(model1$fitted, model1$resid, main="residual plot", xlab="Fitted", ylab="Residuals")
```

Normal transform of rotten tomatoes score.
```{r}
ans <- powerTransform(cbind(mdata$gross, mdata$budget, mdata$imdb_score, t_rt_score-min(t_rt_score)+.01) ~ 1)
summary(ans)
t_gross = (mdata$gross^0.25 - 1)/0.25
t_budget = (mdata$budget^0.25 - 1)/0.25
t_imdb_1 = mdata$imdb_score
t_imdb_2 = (mdata$imdb_score^2 - 1)/2
t_rt = t_rt_score
model2 = lm(t_gross~t_budget+t_imdb_1+t_imdb_2+t_rt)
hexplom(~cbind(t_gross,t_budget,t_imdb_2,t_rt))
summary(model2)
plot(hexbin(model2$fitted, model2$resid), main="Residual Plot", xlab="Fitted", ylab="Residuals", legend=0)

par(mfrow=c(1,1))
plot(hexbin(t_gross, t_budget), main="Gross vs. Budget" , legend=0, xlab="Transformed Gross", ylab="Transformed Budget")
```

## Initial regression
```{r}
rt=mdata
genre_old=rt[,c(9:29)]
Music= as.numeric(genre_old$Music | genre_old$Musical)
genre=cbind(genre_old[,-c(13:14)],Music)

data_initial=rt[,c(2,6,8,10:30)]
data_initial=cbind(data_initial,as.factor(rt$title_year))
lm1=lm(gross~.,data=data_initial)
summary(lm1)
```

## Treat Genres
### Transformation
```{r}
trans1=read.csv("movie_rt_transform.csv")
continious_trans1=trans1[,-c(1:30)]
continious_trans1=cbind(continious_trans1,continious_trans1$t_imdb_score^2,as.factor(rt$title_year))
names(continious_trans1)[c(5,6)]=c("t_imdb_score2","year")
```

### Model selection
```{r}
library(corrplot)
## cluster into 3 groups
corrplot(cor(genre),order="hclust",addrect=3,method="color")
c3_1=as.numeric(genre$Comedy|genre$Romance|genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music)  #8
c3_2=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c3_3=as.numeric(genre$Western|genre$Crime|genre$Thriller|genre$Action|genre$Sci.Fi|genre$Horror|genre$Mystery)   #7
c3=cbind(c3_1,c3_2,c3_3)
colSums(c3)

## cluster into 4 groups
corrplot(cor(genre),order="hclust",addrect=4,method="color")
c4_1=as.numeric(genre$Comedy|genre$Romance)  #2
c4_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music) #6
c4_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c4_4=as.numeric(genre$Western|genre$Crime|genre$Thriller|genre$Action|genre$Sci.Fi|genre$Horror|genre$Mystery)   #7
c4=cbind(c4_1,c4_2,c4_3,c4_4)

## cluster into 5 groups
corrplot(cor(genre),order="hclust",addrect=5,method="color")
c5_1=as.numeric(genre$Comedy|genre$Romance)  #2
c5_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music) #6
c5_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c5_4=as.numeric(genre$Western|genre$Crime|genre$Thriller) #3
c5_5=as.numeric(genre$Action|genre$Sci.Fi|genre$Horror|genre$Mystery)   #7
c5=cbind(c5_1,c5_2,c5_3,c5_4,c5_5)

## cluster into 6 groups
corrplot(cor(genre),order="hclust",addrect=6,method="color")
c6_1=as.numeric(genre$Comedy|genre$Romance)  #2
c6_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music) #6
c6_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c6_4=as.numeric(genre$Western|genre$Crime|genre$Thriller) #3
c6_5=as.numeric(genre$Action|genre$Sci.Fi) #2
c6_6=as.numeric(genre$Horror|genre$Mystery)   #2
c6=cbind(c6_1,c6_2,c6_3,c6_4,c6_5,c6_6)

## cluster into 7 groups
corrplot(cor(genre),order="hclust",addrect=7,method="color")
c7_1=as.numeric(genre$Comedy|genre$Romance)  #2
c7_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family) #6
c7_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c7_4=as.numeric(genre$Western|genre$Crime|genre$Thriller) #3
c7_5=as.numeric(genre$Action|genre$Sci.Fi) #2
c7_6=as.numeric(genre$Horror|genre$Mystery)   #2
c7_7=as.numeric(genre$Documentary|genre$Music)
c7=cbind(c7_1,c7_2,c7_3,c7_4,c7_5,c7_6,c7_7)

data_c3=cbind(continious_trans1,c3,rt$title_year)
lm_c3=lm(t_gross~.,data=as.data.frame(data_c3))
s3=summary(lm_c3)
plot(lm_c3$fitted.values,lm_c3$residuals)

data_c4=cbind(continious_trans1,c4)
lm_c4=lm(t_gross~.,data=as.data.frame(data_c4))
s4=summary(lm_c4)

data_c5=cbind(continious_trans1,c5)
lm_c5=lm(t_gross~.,data=as.data.frame(data_c5))
s5=summary(lm_c5)

data_c6=cbind(continious_trans1,c6)
lm_c6=lm(t_gross~.,data=as.data.frame(data_c6))
s6=summary(lm_c6)

data_c7=cbind(continious_trans1,c7)
lm_c7=lm(t_gross~.,data=as.data.frame(data_c7))
s7=summary(lm_c7)

data_all=cbind(continious_trans1,genre_old)
lm_all=lm(t_gross~.,data=as.data.frame(data_all))
s_all=summary(lm_all)

bic=BIC(lm_c3,lm_c4,lm_c5,lm_c6,lm_c7,lm_all)
```
BIC suggests c6


## Diagnostics
```{r, echo = FALSE}
movie_rt_transform = mdata
continious_trans1=movie_rt_transform[,-c(1:30)]
continious_trans1=cbind(continious_trans1,continious_trans1$t_imdb_score^2,movie_rt_transform[,7])
names(continious_trans1)[5]="t_imdb_score2"

genre_old=movie_rt_transform[,c(9:29)]
Music= as.numeric(genre_old$Music | genre_old$Musical)
genre=cbind(genre_old[,-c(13:14)],Music)

c6_1=as.numeric(genre$Comedy|genre$Romance)  #2
c6_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music) #6
c6_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c6_4=as.numeric(genre$Western|genre$Crime|genre$Thriller) #3
c6_5=as.numeric(genre$Action|genre$Sci.Fi) #2
c6_6=as.numeric(genre$Horror|genre$Mystery)   #2
c6=cbind(c6_1,c6_2,c6_3,c6_4,c6_5,c6_6)

data_c6=cbind(continious_trans1,c6)
lm_c6=lm(t_gross~t_budget+t_imdb_score + t_rt_score +t_imdb_score2 + as.factor(movie_rt_transform[,7])+c6,data=as.data.frame(data_c6))
```

### Influential Points

We calculated cook's distance for the model (shown as below). The maximum cook's distance is around 0.025, which is quite small.

```{r, echo = FALSE}
library(faraway)
library(MASS)
cook <- cooks.distance(lm_c6)
halfnorm(cook,3,labs=movie_rt_transform$movie_title,ylab="Cooks distance")
title(main="Half-Normal Q-Q plot for Cooks distances")
```


### Outliers

The range for standardized residuals is:

```{r, echo = FALSE}
library(faraway)
library(MASS)
range(stdres(lm_c6))
```

### QQ plot for standardized residuals:

```{r, echo = FALSE}
library(faraway)
library(MASS)
qqnorm(stdres(lm_c6), main="Q-Q plot for standardized residuals")
abline(0,1)
```

### Points with large standardized residuals: 

```{r, echo = FALSE}
cbind(fitted(lm_c6),movie_rt_transform[,c("movie_title","t_gross","t_budget","t_imdb_score","t_rt_score")], studres=studres(lm_c6),stdres=stdres(lm_c6))[abs(stdres(lm_c6))>4,]

```

### Dummy variable for Paranormal Activity: (it is significant; this movie is an outlier).

```{r, echo = FALSE}
D1 <- rep(0,length=1329)
names(D1) <- dimnames(movie_rt_transform)[[1]]
D1[1298] <- 1
lm2 <- lm(t_gross~.+D1,data=as.data.frame(data_c6))
summary(lm2) #
```
It turns out to be significant; it is an outlier.

Dummy viriable for Winter's Tale: (it is significant; this movie is an outlier).
```{r, echo = FALSE}
D2 <- rep(0,length=1329)
names(D2) <- dimnames(movie_rt_transform)[[1]]
D2[349] <- 1
lm3 <- lm(t_gross~.+D2,data=as.data.frame(data_c6))
summary(lm3)
```
It turns out to be significant; it is an outlier.

### Model without influential points and outliers:

```{r, echo = FALSE}
lm6=lm(t_gross~.,data=as.data.frame(data_c6),subset=(cook<0.015))
summary(lm6) 
```

## Censor
```{r}
library(censReg)

movie_rt_transform = mdata
continious_trans1=movie_rt_transform[,-c(1:30)]
continious_trans1=cbind(continious_trans1,continious_trans1$t_imdb_score^2,movie_rt_transform[,7])
names(continious_trans1)[5]="t_imdb_score2"
names(continious_trans1)[6]="year"
continious_trans1 = rbind(continious_trans1, 0)
continious_trans1$year = as.factor(continious_trans1$year)

genre_old=movie_rt_transform[,c(9:29)]
Music= as.numeric(genre_old$Music | genre_old$Musical)
genre=cbind(genre_old[,-c(13:14)],Music)

c6_1=as.numeric(genre$Comedy|genre$Romance)  #2
c6_2=as.numeric(genre$Fantasy|genre$Adventure|genre$Animation|genre$Family|genre$Documentary|genre$Music) #6
c6_3=as.numeric(genre$Sport|genre$War|genre$Biography|genre$History) #5
c6_4=as.numeric(genre$Western|genre$Crime|genre$Thriller) #3
c6_5=as.numeric(genre$Action|genre$Sci.Fi) #2
c6_6=as.numeric(genre$Horror|genre$Mystery)   #2
c6=cbind(c6_1,c6_2,c6_3,c6_4,c6_5,c6_6)
c6 = rbind(c6,0)

data_c6=cbind(continious_trans1,c6)

lm_c6=lm(data_c6$t_gross~.,data=as.data.frame(data_c6))
summary(lm_c6)

clm_s6=censReg(formula = data_c6$t_gross~data_c6$t_budget+data_c6$t_imdb_score+data_c6$t_rt_score+
                 data_c6$t_imdb_score2+data_c6$year+data_c6$c6_1+data_c6$c6_2+
                 data_c6$c6_3+data_c6$c6_4+data_c6$c6_5+data_c6$c6_6)
summary(clm_s6)
```
The results for our covariates of interest are very similar.



